% #############################################################################
% Publicações / Publications
% !TEX root = ../main.tex
% #############################################################################
The development of this thesis resulted in several scientific contributions exploring different perspectives of optimization problems:

\begin{enumerate}
\item Caetano, I., Ilunga, G., \textbf{Belém, C.}, Aguiar, R., Feist, S., Bastos, F., and Leitão, A. (2018). \textit{Case Studies on the Integration of Algorithmic Design Processes in Traditional Design Workflows}. Proceedings of the 23rd CAADRIA Conference - Volume 1, Tsinghua University, China, 111-120.

\item \textbf{Belém, C.}, and Leitão, A. (2018). \textit{From Design to Optimized Design An algorithmic-based approach}. Proceedings of the 36th eCAADe Conference - Volume 2, Lodz University of Technology, Poland, 549-558.

\item Martinho, H., Leitão, A, \textbf{Belém, C.}, Loonen, R, and Gomes, M. G. (2019). \textit{Algorithmic Design and Performance Analysis of Adaptive Façades}. Proceedings of the 24th CAADRIA Conference - Volume 1, Victoria University of Wellington, New Zealand, 685-694.

\item \textbf{Belém, C.}, and Leitão, A. (2019). \textit{Conflicting Goals in Architecture: A study on Multi-Objective Optimization}. Proceedings of the 24th CAADRIA Conference - Volume 1, Victoria University of Wellington, New Zealand, 453-462.

\item \textbf{Belém, C.}, Santos, L. S., and Leitão, A. (2019). \textit{On the Impact of Machine Learning: Architecture without Architects?}. Proceedings of the 18th International Conference CAAD Futures 2019, South Korea. (to appear)

\item Pereira, I., \textbf{Belém, C.} and Leitão, A. (2019) \textit{Optimizing Exhibition Spaces: A Multi-Objective Approach}, Proceedings of the 37th eCAADe Conference, Porto University, Portugal (to appear)

\end{enumerate}



