% #############################################################################
% Agradecimentos / Acknowledgments
% !TEX root = ../main.tex
% #############################################################################

I would like to express my respect and gratitude to my supervisor and friend Dr. António Menezes Leitão. He proposed an interesting theme, which proved to be intriguing and challenging. His efforts to arrange research grants and to supply better computational resources were inspiring and encouraged me to fight the difficulties found along the way. His constant support, preoccupation, and first-class guidance were invaluable through this dissertation. Thanks for everything, especially for encouraging me to pursuit my dreams and for providing me with the flexibility and free-will to tackle this theme as something that I would be proud of.  

I would like to thank the members of the research group oriented by my supervisor, Algorithmic Design for Architecture, for their support, valuable ideas, and discussions which undoubtedly improved the practicality of this work - especially, Bruno Ferreira, Inês Caetano, Inês Pereira, Renata Castelo Branco, Guilherme Ilunga, Helena Martinho, and Luís Silveira Santos. 

I would also like to thank the Department of Computer Science and Engineering at \textit{Instituto Superior Técnico}, Universidade de Lisboa for providing me with the foundations for completing this work, as well as for the opportunities to lecture as a teaching assistant during my MSc Thesis. 

I would also like to thank \textit{Fundação para a Ciência e Tecnologia} (FCT) and \textit{Instituto de Engenharia de Sistemas e Computadores - Investigação e Desenvolvimento} (INESC-ID) for the financial support provided in the form of research grants with references UID/CEC/50021/2013, UID/CEC/50021/2019 and PTDC/ART-DAQ/31061/2017.

%I am also grateful to the staff and teachers of the Computer Engineering and Information Systems course for their effort, dedication, availability, and for providing an interesting working environment. 

To all my friends whose support was invaluable during this period and who encouraged me to constantly push my limits when the task felt too large, I thank you from the bottom of my heart - especially, Carolina Pereira, Cristiana Tiago, Diogo Magalhães, Diogo Silva, Filipe Magalhães, Gonçalo Rodrigues, Guilherme Ilunga, Micael Baptista, Nuno Afonso, Pedro Simão, and Telma Correia.

Last but not least, I would like to thank my parents for their friendship, encouragement, and caring over all these years, for always being there for me through thick and thin and without whom this project would not be possible. I would also like to thank my sister, brother, and sister-in-law, for their understanding, support, preoccupation, and love throughout these years.

To each and every one of you -- Thank you.